---
title: "Indexing"
author:
- affiliation: Institute of Medical Biometry and Statistics, Faculty of Medicine and
    Medical Center - University of Freiburg
  name: "Clemens Kreutz, Eva Brombacher, Eva Kohnert, Martin Treppner"
date: "Jan 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(magrittr)
```

## Indexing
```{r}
a <- 1:10
a[1]
a[3:4]
a[a<5]
a[-1]
a[-(1:4)]
a[-(a<5)]
a[-which(a<5)]
a[a>=5]
```

## Assignments
```{r}
a <- 1:10
a[1] <- NA
a
a[a<5] <- NA
a
a[is.na(a)] <- 0
a
```

## Matrices
```{r}
A <- matrix(a,nrow=5)
A
A[,1]
A[A<1]
A[,1]
```

```{r}
t(A[,1])
class(A[,1])
class(A)
t(1:10)
t(t(1:10))
```

## Indexing of lists
```{r}
L <- list(1:4, c(1,4,7),"test")
L
L[1]
L[[1]]
```
```{r}
names(L) <- c("V1","vector2","name")
L
L[["name"]]
L$name
```


## Exceeding the length:
```{r}
a <- 1:10
a[1:13]
```

## Multiplication 
```{r}
a <- 1:5
a
b <- rep(1,times=5)
b
a*b
a %*% b
a %*% t(b)  # an array is a column matrix

```

## Recycling
When applying an operation to two vectors that requires them to be the same length, R automatically recycles, or repeats, elements of the shorter one, until it is long enough to match the longer Vector. 
```{r}
a5 <- 1:5
a5
a2 <- 1:2
a2
b <- rep(2,times=10) 
b

a5 + b
a5 * b
a2 * b
```
## Pipe operator
The following is the same (magrittr package required):
```{r}
a <- 1:10

exp(a)
a %>% exp
```

Pipe operators are helpful to prevent many brackets because of nested functions:

```{r}
round(exp(diff(log(a))), 1)

a %>% log %>% diff %>% exp  %>% round
a %>% log %>% diff %>% exp  %>% round(1)
```


