---
title: "Functions and Scoping"
author:
- affiliation: Institute of Medical Biometry and Statistics, Faculty of Medicine and
    Medical Center - University of Freiburg
  name: "Clemens Kreutz, Eva Brombacher, Eva Kohnert, Martin Treppner"
date: "Jan 2021"
output: html_document
---

## Functions and the scope of variables

A standard example for a function:
```{r }
rm(list=ls())
fun <- function() {
  x <- 1
  y <- 2
  c(x, y)
}
fun()
```

Another example:
```{r}
fun <- function(x) length(unique(x))
fun(c(1,1,1,2,3))
```
Syntax for default arguments:

```{r}
SimData <- function(N=1, mu=0, sigma=1){
  y <- rnorm(N)*sigma + mu
  y
}
SimData()
SimData(10)
SimData(10,sigma=1000)
```


Is x known inside the function?
```{r}
rm(list=ls())
x <- 2
fun <- function() {
  y <- 1
  c(x, y)
}
fun()
```

Is fun1 known inside of fun2?
```{r}
rm(list=ls())
fun1 <- function(x) x + 1
fun2 <- function() {
  fun1(10)
}
fun2()
```

fun1 returns a function.
What's the ouput of fun2()?
```{r}
rm(list=ls())
fun <- function(x) {
  y <- 2
  function() {
    c(x, y)
  }
}
fun2 <- fun(1)
fun2()
```
Now we do exactly the same, but we remove fun before callung fun2.
What happens?

```{r}
rm(list=ls())
fun <- function(x) {
  y <- 2
  function() {
    c(x, y)
  }
}
fun2 <- fun(1)
rm(fun) 
fun2()
```

Nothing. The value y <- 2 is still known !?
This is because during the definition of fun2 the environment is stored:

```{r}
fun2
```

